---
title: "Simulation"
author: "Jungin Choi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning= FALSE}
library(ggplot2)
library(tidyverse)
library(sp)
library(gstat)
library(raster)
library(stats)
```

# Simulation Setting

$$A = {(x_1, x_2) | x_1^2 + x_2^2 \geq 10^2}$$

```{r echo = FALSE, out.width="50%"}
x <- seq(-10,10,length = 50)
y <- seq(-10,10,length = 50)
coords <- expand.grid(x = x, y = y) %>% filter(x^2 + y^2 <= 10^2)

# Plot using ggplot2
ggplot(coords, aes(x = x, y = y)) +
  geom_point(alpha = 0.6, color = 'blue', size = 1) +
  coord_fixed() +
  theme_minimal() +
  labs(title = "Uniformly Distributed Coordinates within a 2D Circular Region",
       x = "X Coordinate",
       y = "Y Coordinate") +
  xlim(-10, 10) +
  ylim(-10, 10) +
  theme(plot.title = element_text(hjust = 0.5))
```

$$\beta_{1,x_1,x_2} = 10|sin(\frac{x_1}{2.5}) + cos(\frac{x_2}{2.5})|$$
$$\beta_{2,x_1,x_2} = 5|sin(\frac{x_1}{3}) + cos(\frac{x_2}{3})|$$

\begin{align*}
  h(t,x_1,x_2) &= \beta_{1,x_1,x_2} \cdot C(t) + \beta_{2,x_1,x_2} \cdot C^{'}(t)\\ 
  n(t) &= 1{onsets}\\
\end{align*}

```{r echo = FALSE, out.width="50%"}
# Example coordinates
set.seed(123)  # For reproducibility
coords <- data.frame(x = coords$x, y= coords$y)
coords$beta1 <- abs(with(coords, sin(x/2.5) * cos(y/2.5)))*10
coords$beta2 <- abs(with(coords, cos(x/3) * sin(y/3)))*5

# Plot the spatial distribution of interpolated values
ggplot(coords, aes(x = x, y = y, fill = beta1)) +
  geom_tile() +
  theme_minimal() +
  labs(title = "Spatially Dependent Variable",
       x = "X Coordinate",
       y = "Y Coordinate",
       fill = "True Beta 1") +
  coord_fixed()

# Plot the spatial distribution of interpolated values
ggplot(coords, aes(x = x, y = y, fill = beta2)) +
  geom_tile() +
  theme_minimal() +
  labs(title = "Spatially Dependent Variable",
       x = "X Coordinate",
       y = "Y Coordinate",
       fill = "True Beta 2") +
  coord_fixed()
```

For simulation, generate 100 timecourse data with AR(p) white noise. In this case, AR(1) model with $\phi = 0.3$.


\begin{align*}
Y(t,x_1,x_2) &= n(t) \cdot h(t,x_1,x_2) + \epsilon(t) & where\ \epsilon \sim AR(1)\\
\epsilon(t) &= \phi \cdot \epsilon(t-1) + \delta, & \delta \sim N(0,\sigma^2)\\
\end{align*}

## 0. Independent Model

$Y_V = n \cdot h + \epsilon  = X\beta_V + \epsilon$ where $\epsilon \sim AR(1)$

where \begin{align*} 
Y = \begin{pmatrix}
Y(1)\\ Y(2)\\ \vdots \\ Y(T) \\
\end{pmatrix}, & X = \begin{pmatrix}
C(1) & C^{'}(1)\\ \vdots \\ C(T) & C^{'}(T) \\
\end{pmatrix}, & \beta = \begin{pmatrix}
\beta_1 \\ \beta_2
\end{pmatrix}
\end{align*}

$\hat{\beta_V} = (X^{'}\Sigma^{-1}X)^{-1}X^{'}\Sigma^{-1}Y_V$ where \begin{align*} 
\Sigma = Cov(\epsilon) = \frac{\sigma^2}{1-\phi^2}\begin{pmatrix}
1 & \phi & \phi^2 & \cdots & \phi^{T-1}\\
\phi & 1 &  & \cdots & \phi^{T-2}\\
\vdots &  &   & \vdots\\
\phi^{T-1} & \cdots &  &  & 1\\
\end{pmatrix}
\end{align*}


## 1. Gaussian Kernel Smoothing

$$\tilde{Y}(t,x_1,x_2) = \frac{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)Y(t,x_{1i}, x_{2i})}{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)}$$

$$\text{where   } K(x_{1}, x_{2}| \mu_1, \mu_2, \sigma) = \frac{1}{2\pi\sigma^2}exp(-\frac{(x_1-\mu_1)^2 + (x_2-\mu_2)^2}{2\sigma^2})$$

\begin{align*} \tilde{Y}_V = \tilde{Y}(x_1,x_2) = \begin{pmatrix}
\tilde{Y}(1,x_1,x_2)\\ \vdots \\ \tilde{Y}(T,x_1,x_2)\\
\end{pmatrix}
\end{align*}

$$\hat{\tilde{\beta_V}} = (X^{'}\Sigma^{-1}X)^{-1}X^{'}\Sigma^{-1}\tilde{Y_V}$$

\begin{align*}\mathbb{E}(\hat{\tilde{\beta_V}}) &= \mathbb{E}((X^{'}\Sigma^{-1}X)^{-1}X^{'}\Sigma^{-1}\tilde{Y_V}) \\
&= (X^{'}\Sigma^{-1}X)^{-1}X^{'}\Sigma^{-1}\mathbb{E}\tilde{Y_V}
\end{align*}

We know that $\mathbb{E}(Y) = X\beta$.

\begin{align*} \mathbb{E}(\tilde{Y}(t,x_1,x_2)) &= \mathbb{E}(\frac{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)Y(t,x_{1i}, x_{2i})}{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)})\\
&= \frac{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)\mathbb{E}(Y(t,x_{1i}, x_{2i}))}{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)}
\end{align*}

$$\mathbb{E}(\tilde{Y}(x_1,x_2)) = \frac{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)X\beta_{x_{1i}, x_{2i}}}{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)}$$

$$\therefore \mathbb{E}(\hat{\tilde{\beta}(x_1,x_2)}) = \frac{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)\beta_{x_{1i}, x_{2i}}}{\sum_i K(x_{1i}, x_{2i}| x_1, x_2, \sigma)}$$


## 2. Thin Plate Spline Regression (TPSR) for Beta

\[Y = XPB\Gamma + \epsilon, \ \ \epsilon \sim N(0,I)\]

where

- $Y = (Y_1' \ \  Y_2' \ \cdots \ \ Y_V')'$: $VT \times 1$ vector
- $X = \begin{pmatrix} 
    X_0 & 0 & \cdots & 0\\
    0 & X_0 & \cdots & 0\\
    \vdots & \ddots & &0\\
    0 &   0   &  \cdots & X_0 
    \end{pmatrix}$: $VT \times VSk_{\beta}$ matrix
- $P =$ Permutation matrix: $VSk_{\beta} \times VSk_{\beta}$ matrix
- $P\begin{pmatrix}
    \beta_{1,1}\\
    \beta_{1,2}\\
    \vdots\\
    \beta_{Sk_{\beta},V-1}\\
    \beta_{Sk_{\beta},V}\\
        \end{pmatrix} = \begin{pmatrix}
    \beta_{1,1}\\
    \beta_{2,1}\\
    \vdots\\
    \beta_{Sk_{\beta}-1,V}\\
    \beta_{Sk_{\beta},V}\\
        \end{pmatrix}$

- $B = \begin{pmatrix}
    B_1 & 0 & \cdots & 0\\
    0 & B_2 & \cdots & 0\\
    \vdots & \ddots & &0\\
    0 &   0     & \cdots& B_{Sk_{\beta}} 
        \end{pmatrix}$: $VSk_{\beta} \times Sk_{\beta}k_{\gamma}$ matrix

- $\Gamma = (\gamma_1' , \cdots, \gamma_{Sk_{\beta}}')'$: $Sk_{\beta}k_{\gamma} \times 1$ vector. 

Then we can fit $\Gamma$ using linear regression model. 

$$\hat{\Gamma} = ((XPB)'(XPB))^{-1}(XPB)'Y$$
Fitted spatial $\beta$ will be expressed as below:

\begin{align*}
\hat{\beta_S} &= \begin{pmatrix}
    \hat{\beta_1}\\
    \hat{\beta_2}\\
    \vdots \\
    \hat{\beta_V}\\
    \end{pmatrix} = PB\hat{\Gamma}
\end{align*}


